---
title: "STA130 Rstudio Demo"
author: "Scott Schwartz"
date: "Fall 2022 Week 1"
output:
  pdf_document: default
---

<!-- Text in here is "html comments" that are not ever printed out -->
<!-- The block up above controls "YAML" settings of this .Rmd file: 
     for example, knitting this .Rmd will result in a pdf document
     
     Markdown provides simple syntax constructions for document formatting   
     Below we make a nice checklist.  For more examples of Markdown, see
     https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf -->

The purpose of this demo is to:

1. [ ] Show how to access Jupyterhub and navigate RStudio
2. [ ] Introduce .Rmd files, html/R comments, and code chunks **[source pane]**
3. [ ] Introduce `parameter=argument` syntax controlling code chunks **[source pane]**
3. [ ] Demonstrate R calculator arithmetic (`+`,`*`,`^`, etc.) and logic (`&`,`|`,`!`, etc.) **[console pane]**
4. [ ] Discuss loadable packages like `readr`, `tibble`, and `tidyverse` which add   
   functions like `readr::read_csv()` and `tibble::tibble()` to the R calculator
5. [ ] Introduce `<-`, `->`, and `=` variable assignments in R **[source/console pane]**
6. [ ] Show variable and data set management and the `rm()` function **[environment pane]**
7. [ ] Introduce the `<cmd-shift-M>` (Mac) and `<ctrl-shift-M>` (PC) "pipe symbol" `%>%`  
8. [ ] Show how to run code with `<cmd-shift-return>` (Mac) and `<ctrl-shift-enter>` (PC),  
   and the "play" and "knit" buttons **[source pane]**
9. [ ] Show exporting/downloading files from RStudio, organizing and managing files,  
   and course Rmd+pdf submission protocols **[files pane]**
10. [ ] Highlight Tutorial, Packages, and Help tabs **[environment/files pane]**


<!-- Below is code chunk where R code goes -->
<!-- The `parameter=argument` configuration in the curly brackets controls 
     how the code chunk is used during knitting -->
```{r, eval=TRUE, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Code chunks have `#` "R comments" instead of `<!-- -->` "html comments"
# The `eval=TRUE` means this code will be run during knitting
# The `include=FALSE` means neither code nor output will be included in the knit
# The `echo=FALSE` (with `include=TRUE`) will print the output but not code chunk  
# The `message=FALSE` means informational messages won't be shown in the knit
# The `warning=FALSE` means warning alert message won't be shown in the knit
# https://rmarkdown.rstudio.com/lesson-3.html

# Rather than always specifying these, we'll set them to all default to `FALSE`
# This means knitting completely ignore all code-chunk by default
knitr::opts_chunk$set(eval=FALSE, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE)
# `knitr::` means not having to run `library(knitr)` to access `opts_chunk`
```

<!-- Knitting will completely ignore this code chunk under the above defaults -->
```{r} 
# load tidyverse functionality <-- make code "easy to understand" with comments!
library(tidyverse) # `install.packages("tidyverse")` isn't needed 
                   #  on jupyterhub for the `tidyverse` package
```

## Question 1

Run the R code chunk below that contains `tidyverse` functions to load and show a data set of coffee ratings.

- use key-binding shortcuts cmd-shift-return (Mac) or ctrl-shift-enter (PC), or the green "play" button 
- "could not find function" for `read_csv()` or `glimpse()` is fixed by running `library(tidyverse)` 
- "cannot open file" is fixed by putting the file with the right name in the same directory as the .Rmd


```{r} 
coffee_ratings <- read_csv("coffee_ratings.csv") # rather than `read.csv()`
head(coffee_ratings)        # <- compare what this line of code does
#coffee_ratings %>% head() # <- versus what this line of code does
# The `%>%` "pipe symbol" is created with <ctrl-shift-M> or <cmd-shift-M>
```
<!-- The "pipe symbol" in other context is sometimes `|`; but, since this is
     the logical `or` operator in R, for R the "pipe symbol is `%>%` -->


### (a) Use the glimpse() function output above to determine how many rows and columns there are in the coffee ratings dataset. 

```{r} 
glimpse(coffee_ratings)        # <- compare what this line of code does
# coffee_ratings %>% glimpse() # <- versus what this line of code does
# The `%>%` "pipe symbol" is created with <ctrl-shift-M> or <cmd-shift-M>
```

### (b) Change the property of the R code chunks above so the knit file shows the code and both the message from `read_csv` and the output from `glimpse`.

### (c) Change the property of the R code chunk below so the plot of overall coffee ratings appears in the knitted output.

```{r}
# Can you read this code? What do you roughly think it's doing?
coffee_ratings %>% # E.g., what does the `%>%` seem to be doing?
  ggplot(aes(x=total_cup_points)) + # E.g., what about `+` !??
  geom_histogram(bins=30, color="black", fill="light blue")
  # Note here that R code uses the same `parameter=argument` pairing as knitting
```

## Module 1 Homework Preview

```{r}
# https://www.tidyverse.org/
# `install.packages("tidyverse")` isn't needed on jupyterhub for `tidyverse`
library(tidyverse)
```

```{r}
# https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-08-11/readme.md
#install.packages("appa")
# https://github.com/averyrobbins1/appa
#devtools::install_github("averyrobbins1/appa")
library(appa)
?appa
```

```{r, include=TRUE}
avatar <- appa # read_csv("avatar.csv")
avatar %>% glimpse()
```

## More Resources

Looking for other presentations of this material? Try out the following resources.

- [R for Data Science Chapter 4: Workflow basics](https://r4ds.had.co.nz/workflow-basics.html)

- [Rstudio Primer: Programming Basics](https://rstudio.cloud/learn/primers/1.2)

- [The Department of Statistical Science Toolkit](https://dosstoolkit.com/)

